---
title: "Project Writeup - Practical Machine Learning"
output: html_document
---

#Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement  a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)."

#Data Cleaning

Data import, importing Blanks, "NA" and "#DIV/0!" as NA's.
```{r}
setwd("/Users/Sheetanshu/Coursera_R/ML")
pmlTrain<-read.csv("pml-training.csv", header=T, na.strings=c("","NA", "#DIV/0!"))
pmlTest<-read.csv("pml-testing.csv", header=T, na.string=c("","NA", "#DIV/0!"))
```

Pre-proessing the training data set.Removing the variables with at least one "NA" from the analysis. We are now left with 60 Variables instead of 160.
```{r}
noNATrain<-pmlTrain[, apply(pmlTrain, 2, function(x) !any(is.na(x)))] 
dim(noNATrain)
```

Also removing the variables which were related to time and user information. Now we have 52 variables.
```{r}
cleanTrain<-noNATrain[,-c(1:8)]
dim(cleanTrain)
```

Same variables were retained in Test data set (Validation Data set) to be used for predicting the test 20 cases.
```{r}
cleantest<-pmlTest[,names(cleanTrain[,-52])]
dim(cleantest)
```

#Data Partition and Process

Partitioning the cleaned Training data set into 70% Training and 30% Test data sets.
We would be using **Random Forests** Technique to predict the outcome of this exercise.

```{r}
library(caret)
inTrain<-createDataPartition(y=cleanTrain$classe, p=0.70,list=F)
training<-cleanTrain[inTrain,] 
test<-cleanTrain[-inTrain,] 
#Training and test set dimensions
dim(training)
dim(test)
```

#Algorithm

**Random Forest** trees were generated for the 70% Training data set
using cross-validation. 

```{r}
library(caret)
set.seed(1234)
#Build trainControl for Cross - Validation
TC<-trainControl(method="cv", number=5, allowParallel=T, verbose=T)
TreeFit<-train(classe~.,data=training, method="rf", trControl=TC, verbose=F)
```

The algorithm was checked on the paritioned 30% Test data set to calculate the accuracy and estimated error of prediction.
```{r}
TestPred<-predict(TreeFit,newdata=test)
confusionMatrix(TestPred,test$classe)
```

#Project Submission

Predicting the output for 20 cases provided in the validation data set.
```{r}
Val20<-predict(TreeFit,newdata=cleantest)
Val20
```

Writing 20 Files using the script provided.
```{r}
setwd("/Users/Sheetanshu/Coursera_R/ML")
getwd()
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(Val20)
```

Other Algorithms were also tested but Random Forest yields the best accuracy. The prediction for the 20 cases were also similar.

#Conclusion

**Out of Sample Error Estimation**
```{r}
outOfSampleError.accuracy<-sum(TestPred==test$classe)/length(TestPred)
outOfSampleError <- (1 - outOfSampleError.accuracy)*100
paste0("Out of sample error estimation: ", round(outOfSampleError,digits=2),"%")
```
The out of sample error estimation is 0.66%


**Accuracy**
```{r}
paste0("Confusion Matrix accuracy estimation: ", round((outOfSampleError.accuracy)*100,digits=2),"%")
```

The Accuracy estimation is 99.34%